Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-05-24 docs/Manuscript.Rmd
Close #3 : Added changes to MnM

Governments and non-profit organizations have sought to reduce lionfish densities through removal programs and incentivizing its consumption \citep{chin_2016}. In some cases, these have shown to significantly reduce --but not quite eliminate-- lionfish abundances at local scales \citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs can help stabilize or grow native prey fish populations \citep{cote_2014}. Complete eradication of lionfish through fishing is unlikely because of their rapid recovery rates and ongoing recruitment to shallow-water areas from their persistent populations in mesophotic coral ecosystems \citep{barbour_2011,andradibrown_2017}. However, promoting its consumption might create a level of demand capable of sustaining a stable fishery, which can help control invasive populations while providing alternative livelihoods \citep{chin_2016}.

@@

Empirical modeling efforts examining the feasibility of establishing fisheries for lionfish involve modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via age- or length-structured population models \citep{cote_2014,barbour_2011,andradibrown_2017}, which convert fish length or age to weight to then calculate total biomass. Therefore, length-weight relationships are an essential component of biomass-based modeling. However, this relationship can vary across regions as a response to biotic (*e.g.* local food availability) and abiotic (*e.g.* water temperature) conditions \citep{johnson_2016}. Literature suggests that when length-weight relationships are spatially variable, Site-specific parameters are necessary to obtain accurate biomass estimates. This becomes ingreasingly important when estimating the potential effectiveness (and resources neded) of lionfish culling programs or identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. Genetic analysis of invasive lionfish suggest that two distinct population groups exist \citep{betancurr_2011}, and as such we would expect their biology to differ. Despite the large number of studies reporting site-specific length-weight relationships, no studies have described region-wide differences in these.

@@

The main objective of this work was to compare allometric growth of lionfish throughout their invasion range. Allometric parameters were retreived from scientific literature, and an additional pair of parameters was calculated from filed observations in the central Mexican Caribbean.

Length-weight relationships (n = `r dim(review)[1]-1`) identified in literature were obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and Caribbean (n = 10). Information on sampling methods, sex differentiation, location, and depth ranges of each study was retrieved when available. Studies were assumed to include both genders if gender was unspecified. Locations where allometric studies were performed are shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

Studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept ($c$) from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. Here, all parameters are reported as TL(mm) to TW(gr) conversions. When required, values from other studies were transformed for consistency.

<!-- Make this part more clear. Explain that the goal is to show "how far off" the estimates would be -->

We compare observed TW to TW estimated with the 18 length-weight relationsips to quantify the extent to which using *ex situ* this leads to over- or underestimation. Predicted weights were divided by the observed weights to obtain a ratio. Difference in mean weight ratios across studies were tested with a one-way Analysis of Variance (ANOVA). All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available at dryad.org.

>>>>

Governments and non-profit organizations have sought to reduce lionfish densities through removal programs and incentivizing its consumption \citep{chin_2016}. In some cases, these have shown to significantly reduce --but not quite eliminate-- lionfish abundances at local scales \citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs can help stabilize or grow native prey fish populations \citep{cote_2014}. Complete eradication of lionfish through fishing is unlikely because of their rapid recovery rates and ongoing recruitment to shallow-water areas from their persistent populations in mesophotic coral ecosystems \citep{barbour_2011,andradibrown_2017}. However, promoting its consumption might create a level of demand capable of sustaining a stable fishery, which can help control shallow-water populations while providing alternative livelihoods and avoiding further impacts to local reef biota \citep{chin_2016}.

@@

Empirical modeling efforts examining the feasibility of establishing fisheries for lionfish involve modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via age- or length-structured population models \citep{cote_2014,barbour_2011,andradibrown_2017}, which convert fish length or age to weight to then calculate total biomass. Therefore, length-weight relationships are an essential component of these models. However, the length-weight relationship can vary across regions as a response to biotic (*e.g.* local food availability) and abiotic (*e.g.* water temperature) conditions \citep{johnson_2016}. Literature suggests that site-specific parameters are necessary to obtain accurate biomass estimates when length-weight relationships present spatial variation. This becomes ingreasingly important when estimating the potential effectiveness (and resources neded) of lionfish culling programs or identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. Genetic analysis of invasive lionfish suggest that two distinct population groups exist \citep{betancurr_2011}, and as such we would expect their biology to differ. Despite the large number of studies reporting site-specific length-weight relationships, no studies have described region-wide differences in these.

@@

The main objective of this work was to compare allometric growth of lionfish throughout their invasion range. Allometric parameters were retreived from scientific literature, and an additional pair of parameters were calculated from filed observations in the central Mexican Caribbean.

Length-weight relationships (n = `r dim(review)[1]-1`) identified in literature were obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and Caribbean (n = 10). Information on sampling methods, sex differentiation, location, and depth ranges of each study was retrieved when available. Studies were assumed to include both genders if gender was unspecified. Studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept ($c$) from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. Here, all parameters are reported as TL(mm) to TW(gr) conversions. When required, values from other studies were transformed for consistency. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

<!-- Still having trouble to explain this paragraph. Perhaps you can make it better? The idea is that we test wether the differences are trivial by taking all 109 lionfish and calculating TW with each of the 18 available relationships. We then take those, and divide them by the true weights, and obtain a ratio that is easy to interpret if theres an over or under estimation... andwe do some bassic stats on them -->

Combining the length-weight parameters extracted from the literature and the additional pair calculated here, we obtain a total of `r dim(review)[1]` pairs. We test wether using *ex situ* parameters produces differences in weight estimates by estimating TW for the 109 length observations of lionfish from the central Mexican Caribbean with each of the `r dim(review)[1]` relationships. Predicted weights were divided by know observed weights to obtain a simple measurement of over- or underestimation. Difference in mean weight ratios across studies were tested with a one-way analysis of variance (ANOVA). All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available at dryad.org.

### 2018-05-24 docs/Manuscript.tex
Close #3 : Added changes to MnM

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling). A common way to model
this is via age- or length-structured population models
\citep{cote_2014,barbour_2011,andradibrown_2017}, which convert fish
length or age to weight to then calculate total biomass. Therefore,
length-weight relationships are an essential component of biomass-based
modeling. However, this relationship can vary across regions as a
response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}.
Literature suggests that when length-weight relationships are spatially
variable, Site-specific parameters are necessary to obtain accurate
biomass estimates. This becomes ingreasingly important when estimating
the potential effectiveness (and resources neded) of lionfish culling
programs or identifying total biomass available for harvest by fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}.
Genetic analysis of invasive lionfish suggest that two distinct
population groups exist \citep{betancurr_2011}, and as such we would
expect their biology to differ. Despite the large number of studies
reporting site-specific length-weight relationships, no studies have
described region-wide differences in these.

@@

Length-weight relationships (n = 17) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 10). Information on sampling methods, sex
differentiation, location, and depth ranges of each study was retrieved
when available. Studies were assumed to include both genders if gender
was unspecified. Locations where allometric studies were performed are
shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

We compare observed TW to TW estimated with the 18 length-weight
relationsips to quantify the extent to which using \emph{ex situ} this
leads to over- or underestimation. Predicted weights were divided by the
observed weights to obtain a ratio. Difference in mean weight ratios
across studies were tested with a one-way Analysis of Variance (ANOVA).
All analyses were performed in R version 3.5.0 \citep{rcore_2018}. Raw
data and code used in this work are available at dryad.org.

>>>>

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling). A common way to model
this is via age- or length-structured population models
\citep{cote_2014,barbour_2011,andradibrown_2017}, which convert fish
length or age to weight to then calculate total biomass. Therefore,
length-weight relationships are an essential component of these models.
However, the length-weight relationship can vary across regions as a
response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}.
Literature suggests that site-specific parameters are necessary to
obtain accurate biomass estimates when length-weight relationships
present spatial variation. This becomes ingreasingly important when
estimating the potential effectiveness (and resources neded) of lionfish
culling programs or identifying total biomass available for harvest by
fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}.
Genetic analysis of invasive lionfish suggest that two distinct
population groups exist \citep{betancurr_2011}, and as such we would
expect their biology to differ. Despite the large number of studies
reporting site-specific length-weight relationships, no studies have
described region-wide differences in these.

@@

Length-weight relationships (n = 17) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 10). Information on sampling methods, sex
differentiation, location, and depth ranges of each study was retrieved
when available. Studies were assumed to include both genders if gender
was unspecified. Locations where allometric studies have been performed
are shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

Combining the length-weight parameters extracted from the literature and
the additional pair calculated here, we obtain a total of 18 pairs. We
test wether using \emph{ex situ} parameters produces differences in
weight estimates by estimating TW from the 109 length observations of
lionfish from the central Mexican Caribbean with each ot the 18
relationships. Predicted weights were divided by know observed weights
to obtain a simple measurement of over- or underestimation. Difference
in mean weight ratios across studies were tested with a one-way analysis
of variance (ANOVA). All analyses were performed in R version 3.5.0
\citep{rcore_2018}. Raw data and code used in this work are available at
dryad.org.
